---
title: "01_make_data_miriam.Rmd"
author: "Miriam Holleran-Meyer"
output: html_notebook
---

Section 1: Describe the Project

The Trash Wheel Collection Data tracks the amount of trash from dumpsters over time. It keeps track of the weight and volume of the trash, as well as the amounts and types of trash collected. Additionally, it tells how many homes are powered from the the dumpster's trash.



Section 2: What you hope to discover in your killer graph

I hope to be able to see the relationships between the trash collected and the homes it powers, and how these relationships may vary over time.



Section 3: Code and text to retrieve the dataset

Loading trash data:
```{r}
library(tidyverse)
library(tidytuesdayR)
library(ggplot2)
tuesdata <- tidytuesdayR::tt_load(2024, week = 10)
trashwheel <- tuesdata$trashwheel
```



Section 4: Verify dimensions and variables

First lines of trash data:
```{r}
head(trashwheel)
```


Section 5: Save locally to sub-folder "RData"

```{r}
# creating RData
dir.create("RData")

# Save the data to the sub-folder
save(trashwheel, file = "RData/trashwheel.RData")
```



