---
title: "Exploring Trash Wheel Collection Data"
author: "Kunal Kumar"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(here)
library(ggplot2) 
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction to Data

The Trash Wheel Collection Data is part of the Healthy Harbor initiative in Baltimore, which uses semi-autonomous trash interceptors to remove trash from waterways. This dataset provides insights into the types and quantities of trash collected, offering the ability to explore patterns, seasonal variations, and the impact of these efforts on environmental sustainability.

## Data Cleaning and Exploration

## Loading Data

```{r load-data}
# load the data directly from GitHub.
trashwheel <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2024/2024-03-05/trashwheel.csv')
summary(trashwheel)
glimpse(trashwheel)
```
### Data Cleaning

```{r cleaning-data}
# ensure consistent column names and remove unnecessary columns
trashwheel <- trashwheel %>% 
  rename(
    Weight = Weight,
    Volume = Volume,
    PlasticBottles = PlasticBottles,
    Polystyrene = Polystyrene,
    CigaretteButts = CigaretteButts,
    GlassBottles = GlassBottles,
    PlasticBags = PlasticBags,
    Wrappers = Wrappers,
    SportsBalls = SportsBalls,
    HomesPowered = HomesPowered
  ) %>% 
  filter(!is.na(Weight))

glimpse(trashwheel)
```

### Summary Statistics

```{r summary-statistics}
# summarize key statistics
summary_stats <- trashwheel %>% 
  summarize(
    TotalWeight = sum(Weight, na.rm = TRUE),
    TotalPlasticBottles = sum(PlasticBottles, na.rm = TRUE),
    TotalPolystyrene = sum(Polystyrene, na.rm = TRUE),
    TotalCigaretteButts = sum(CigaretteButts, na.rm = TRUE)
  )

summary_stats
```

# Why This Data Matters

The Trash Wheel Collection Data provides crucial insights into the effectiveness of environmental initiatives aimed at reducing waterway pollution. By quantifying the types and amounts of trash collected, this data highlights the scale of waste entering our ecosystems and underscores the importance of sustainable waste management practices. Moreover, analyzing seasonal and regional variations can inform targeted interventions and public awareness campaigns, driving broader participation in environmental conservation efforts. This is the team consensus document.