---
title: "Exploring Trash Wheel Collection Data"
author: "Kunal Kumar"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(here)
library(ggplot2) 
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

The Trash Wheel Collection Data is part of the Healthy Harbor initiative in Baltimore, which uses semi-autonomous trash interceptors to remove trash from waterways. This dataset provides insights into the types and quantities of trash collected, offering opportunities to explore patterns, seasonal variations, and the impact of these efforts on environmental sustainability.

This analysis will clean and explore the dataset, focusing on key questions such as:

- What types of trash are most commonly collected?
- Do different Trash Wheels collect different types of trash?
- Are there seasonal trends in trash collection?

## Data Cleaning and Exploration

### Loading Data

```{r load-data}
# load the data directly from GitHub.
trashwheel <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2024/2024-03-05/trashwheel.csv')
summary((trashwheel))
glimpse(trashwheel)
```

### Data Cleaning

```{r cleaning-data}
# ensure consistent column names and remove unnecessary columns
trashwheel <- trashwheel %>% 
  rename(
    Weight = Weight,
    Volume = Volume,
    PlasticBottles = PlasticBottles,
    Polystyrene = Polystyrene,
    CigaretteButts = CigaretteButts,
    GlassBottles = GlassBottles,
    PlasticBags = PlasticBags,
    Wrappers = Wrappers,
    SportsBalls = SportsBalls,
    HomesPowered = HomesPowered
  ) %>% 
  filter(!is.na(Weight))

glimpse(trashwheel)
```

### Summary Statistics

```{r summary-statistics}
# summarize key statistics
summary_stats <- trashwheel %>% 
  summarize(
    TotalWeight = sum(Weight, na.rm = TRUE),
    TotalPlasticBottles = sum(PlasticBottles, na.rm = TRUE),
    TotalPolystyrene = sum(Polystyrene, na.rm = TRUE),
    TotalCigaretteButts = sum(CigaretteButts, na.rm = TRUE)
  )

summary_stats
```

### Distribution of Trash Types

```{r trash-types}
# visualize the distribution of different types of trash
trash_types <- trashwheel %>% 
  select(PlasticBottles, Polystyrene, CigaretteButts, GlassBottles, PlasticBags, Wrappers, SportsBalls) %>% 
  pivot_longer(cols = everything(), names_to = "TrashType", values_to = "Count") %>% 
  group_by(TrashType) %>% 
  summarize(TotalCount = sum(Count, na.rm = TRUE))

# plot
ggplot(trash_types, aes(x = TrashType, y = TotalCount, fill = TrashType)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  labs(title = "Distribution of Trash Types", x = "Type of Trash", y = "Total Count")
```

### Seasonal Trends

```{r seasonal-trends}
# analyze seasonal trends in trash collection
trashwheel <- trashwheel %>% 
  mutate(Date = as.Date(Date, format = "%Y-%m-%d"),
         Month = factor(Month, levels = month.name))

seasonal_trends <- trashwheel %>% 
  group_by(Month) %>% 
  summarize(TotalWeight = sum(Weight, na.rm = TRUE))

# plot
ggplot(seasonal_trends, aes(x = Month, y = TotalWeight, group = 1)) +
  geom_line() +
  geom_point() +
  theme_minimal() +
  labs(title = "Seasonal Trends in Trash Collection", x = "Month", y = "Total Weight (tons)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

# Why This Data Matters

The Trash Wheel Collection Data provides crucial insights into the effectiveness of environmental initiatives aimed at reducing waterway pollution. By quantifying the types and amounts of trash collected, this data highlights the scale of waste entering our ecosystems and underscores the importance of sustainable waste management practices. Moreover, analyzing seasonal and regional variations can inform targeted interventions and public awareness campaigns, driving broader participation in environmental conservation efforts.

This dataset is not just a snapshot of trash collection—it is a testament to the power of innovative solutions like Mr. Trash Wheel in addressing global challenges like pollution and climate change.
