---
title: "02_EDA_miriam"
author: "Miriam Holleran-Meyer"
output: html_document
---


```{r setup, include=FALSE}
library(ggcorrplot)
library(tidyverse)
library(here)
library(ggplot2) 
knitr::opts_chunk$set(echo = TRUE)
library(tidytuesdayR)
tuesdata <- tidytuesdayR::tt_load(2024, week = 10)
trashwheel <- tuesdata$trashwheel
```



Section 1: Identify response variables, Examine distributions

Summary Statistics for Numerical Variables:
```{r}
summary(trashwheel[c("Weight", "Volume", "PlasticBottles", "Polystyrene", "CigaretteButts", "GlassBottles", "PlasticBags", "Wrappers", "SportsBalls", "HomesPowered")])
```

Distributions of Potential Response Variables:
```{r}
trashwheel %>% 
  ggplot(aes(x = Weight)) + 
  geom_histogram(binwidth = 0.25, fill = "blue", color = "black") + 
  labs(title = "Distribution of Weight", x = "Weight", y = "Frequency")

trashwheel %>% 
  ggplot(aes(x = Volume)) + 
  geom_histogram(binwidth = 1, fill = "blue", color = "black") + 
  labs(title = "Distribution of Volume", x = "Volume", y = "Frequency")

trashwheel %>% 
  ggplot(aes(x = HomesPowered)) + 
  geom_histogram(binwidth = 2, fill = "blue", color = "black") + 
  labs(title = "Distribution of HomesPowered", x = "Volume", y = "Frequency")
```



Section 2: Identify explanatory variables, Make histograms or bar graphs, Look for imbalance

Distributions of Potential Explanatory Variables:
```{r}
trashwheel %>% 
  ggplot(aes(x = Weight)) + 
  geom_histogram(binwidth = 0.25, fill = "blue", color = "black") + 
  labs(title = "Distribution of Weight", x = "Weight", y = "Frequency")

trashwheel %>% 
  ggplot(aes(x = PlasticBottles)) + 
  geom_histogram(binwidth = 400, fill = "blue", color = "black") + 
  labs(title = "Distribution of Plastic Bottles", x = "Plastic Bottles", y = "Frequency")

trashwheel %>% 
  ggplot(aes(x = Polystyrene)) + 
  geom_histogram(binwidth = 400, fill = "blue", color = "black") + 
  labs(title = "Distribution of Polystyrene", x = "Polystyrene", y = "Frequency")

trashwheel %>% 
  ggplot(aes(x = CigaretteButts)) + 
  geom_histogram(binwidth = 20000, fill = "blue", color = "black") + 
  labs(title = "Distribution of Cigarette Butts", x = "Cigarette Butts", y = "Frequency")

trashwheel %>% 
  ggplot(aes(x = GlassBottles)) + 
  geom_histogram(binwidth = 5, fill = "blue", color = "black") + 
  labs(title = "Distribution of Glass Bottles", x = "Glass Bottles", y = "Frequency")

trashwheel %>% 
  ggplot(aes(x = PlasticBags)) + 
  geom_histogram(binwidth = 1000, fill = "blue", color = "black") + 
  labs(title = "Distribution of Plastic Bags", x = "Plastic Bags", y = "Frequency")

trashwheel %>% 
  ggplot(aes(x = Wrappers)) + 
  geom_histogram(binwidth = 1000, fill = "blue", color = "black") + 
  labs(title = "Distribution of Wrappers", x = "Wrappers", y = "Frequency")

trashwheel %>% 
  ggplot(aes(x = SportsBalls)) + 
  geom_histogram(binwidth = 2, fill = "blue", color = "black") + 
  labs(title = "Distribution of Sports Balls", x = "Sports Balls", y = "Frequency")
```


Section 3: Review goal of analysis, Make preliminary graph

Goal of analysis: looking at relationships between the trash collected and the homes it powers, and how these relationships may vary over time

Preliminary Graphs: (taken from Kunal's Code)

Showing distribution of trash types:
```{r}
# visualize the distribution of different types of trash
trash_types <- trashwheel %>% 
  select(PlasticBottles, Polystyrene, CigaretteButts, GlassBottles, PlasticBags, Wrappers, SportsBalls) %>% 
  pivot_longer(cols = everything(), names_to = "TrashType", values_to = "Count") %>% 
  group_by(TrashType) %>% 
  summarize(TotalCount = sum(Count, na.rm = TRUE))

# plot
ggplot(trash_types, aes(x = TrashType, y = TotalCount, fill = TrashType)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  labs(title = "Distribution of Trash Types", x = "Type of Trash", y = "Total Count")
```

Looking at seaosnal trends:
```{r}
# analyze seasonal trends in trash collection
trashwheel <- trashwheel %>% 
  mutate(Date = as.Date(Date, format = "%Y-%m-%d"),
         Month = factor(Month, levels = month.name))

seasonal_trends <- trashwheel %>% 
  group_by(Month) %>% 
  summarize(TotalWeight = sum(Weight, na.rm = TRUE)) %>%
  filter(!is.na(TotalWeight) & TotalWeight > 0)

# plot
ggplot(seasonal_trends, aes(x = Month, y = TotalWeight, group = 1)) +
  geom_line() +
  geom_point() +
  theme_minimal() +
  labs(title = "Seasonal Trends in Trash Collection", x = "Month", y = "Total Weight (tons)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```



Section 4: Save a new dataset, if necessary

Data Cleaning: (used from Kunal's Code)
```{r}
# ensure consistent column names and remove unnecessary columns
trashwheel <- trashwheel %>% 
  rename(
    Weight = Weight,
    Volume = Volume,
    PlasticBottles = PlasticBottles,
    Polystyrene = Polystyrene,
    CigaretteButts = CigaretteButts,
    GlassBottles = GlassBottles,
    PlasticBags = PlasticBags,
    Wrappers = Wrappers,
    SportsBalls = SportsBalls,
    HomesPowered = HomesPowered
  ) %>% 
  filter(!is.na(Weight))

glimpse(trashwheel)
```

Saving new data:
```{r}
# creating RData
dir.create("RData")

# Save the data to the sub-folder
save(trashwheel, file = "RData/trashwheel.RData")
```


